Lazy Linearity for a Core Functional Language

### 2025-10-17 popl26.tex
reconcile with -> accept

\item We introduce Linear Core, a \ROUNDTWO{lazy} linear language with
the key features of Core (except for type equality coercions), whose type
system \ROUNDTWO{is designed to reconcile with the lazy} semantics of
linearity. \ROUNDTWO{Our system achieves this via a novel combination
  of features: explicit tracking linear dependencies of variables
  through \emph{usage environments}
  (Section~\ref{sec:usage-environments}), codifying that lazy variable
  bindings do not consume resources upon definition, but rather when
  the variables are consumed;
  a distinct treatment of case
  scrutinees in weak-head normal form (Section~\ref{sec:onwhnf}),
  identifying that linear resources captured by case expressions in
  WHNF can be safely used in branches since the case will not further
  evaluate its scrutinee; and a notion of irrelevant and tagged resources
  (Section~\ref{sec:irrev}), which allow for fine-grained control over
  variable usage. Namely, linear resources occurring in non-WHNF case
  scrutinees can no longer be used in the alternative, while the
  pattern variables (or the case binder) must necessarily be used. } 
%
To the best of our
knowledge, this is the first type system \ROUNDTWO{to accept programs
which depend on non-strict evaluation to be understood as linear}
(Section~\ref{sec:main:linear-core}).

>>>>

\item We introduce Linear Core, a \ROUNDTWO{lazy} linear language with
the key features of Core (except for type equality coercions), whose type
system \ROUNDTWO{is designed to accept the lazy} semantics of
linearity. \ROUNDTWO{Our system achieves this via a novel combination
  of features: explicit tracking linear dependencies of variables
  through \emph{usage environments}
  (Section~\ref{sec:usage-environments}), codifying that lazy variable
  bindings do not consume resources upon definition, but rather when
  the variables are consumed;
  a distinct treatment of case
  scrutinees in weak-head normal form (Section~\ref{sec:onwhnf}),
  identifying that linear resources captured by case expressions in
  WHNF can be safely used in branches since the case will not further
  evaluate its scrutinee; and a notion of irrelevant and tagged resources
  (Section~\ref{sec:irrev}), which allow for fine-grained control over
  variable usage. Namely, linear resources occurring in non-WHNF case
  scrutinees can no longer be used in the alternative, while the
  pattern variables (or the case binder) must necessarily be used. } 
%
To the best of our
knowledge, this is the first type system \ROUNDTWO{to accept programs
which depend on non-strict evaluation to be understood as linear}
(Section~\ref{sec:main:linear-core}).
